---
title: "Create Randomised Pretest Questions - Final version"
author: "Stefanie Meliss"
date: "`r Sys.Date()`"
output: html_document
---


Based upon the results of the pretest, some items and/or distractors were modified, please see change log below. Items and/or distractors were modified if the correct response rate in the pilot study was ≥ 80%. In cases where distractor edits were not possible and items needed to be modified, the scope of the item and thereby the breadth of the prior knowledge assessment remained unchanged. The wording of items was also modified in response to participant feedback to avoid ambiguity.  
  
**Q4** [A grapheme is an individual letter or a group of letters that represents…]  
Distractor "A word" was changed to "The sound of a whole word"  
  
**Q7**original question: "Reading fluency contributes to reading comprehension by… " changed to "Which of these capabilities is least likely to contribute to a person’s reading fluency?" together with all distractors  
  
**Q8** original question "Tier 3 vocabulary can be thought of as words that…" changed to "Vocabulary tiers form a construct that is commonly used to prioritise particular words to teach. Tier 3 vocabulary can be thought of as…" together with all distractors  
  
**Q9** original question "Which of these comprehension strategies does not have a sound basis in evidence?" reworded to "Which of these comprehension strategies has the least sound basis in evidence?"  
  
**Q10** original question "Pupils make inferences when they read…" changed to "In terms of reading, inferences are conclusions that…" together with all distractors  
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```

```{r, include=FALSE, echo=F}
# ---------- define all questions and corresponding answers ----------

# q1
q01_q  <- "Reading comprehension can be described as relying on two capacities: the ability to recognise words and build meaning from words. These are generally described as..."
q01_c  <- "Word recognition and language comprehension."
q01_d1 <- "Word recognition and fluency." 
q01_d2 <- "Word recognition and reading comprehension." 
q01_d3 <- "Word recognition and background knowledge."  

# q2
q02_q  <- "An orthography is..." 
q02_d1 <- "The set of conventions associated with a phonics programme." 
q02_d2 <- "The set of conventions associated with a spoken language." 
q02_c  <- "The set of conventions associated with a written language." 
q02_d3 <- "The set of conventions associated with a handwriting system."

# q3
q03_q <- "A phoneme can be thought of as..." 
q03_c  <- "The smallest chunk of spoken sound that can distinguish one word from another."
q03_d1 <- "Any chunk of sound that is smaller than a whole word."
q03_d2 <- "The chunk of sound associated with a syllable."
q03_d3 <- "The chunk of sound associated with a morpheme." 

# q4
q04_q    <- "A grapheme is an individual letter or a group of letters that represents... "
q04_d1 <- "A syllable."
q04_d2 <- "A morpheme." 
q04_d3 <- "The sound of a whole word."
q04_c  <- "A phoneme." 

# q5
q05_q  <- "Alongside chunks of spoken sound, the English writing system represents the smallest chunks within words that are associated with meaning called..."
q05_d1 <- "Syntactic identifiers."
q05_c  <- "Morphemes."
q05_d2 <- "Lexemes."
q05_d3 <- "Graphemes."

# q6
q06_q  <- "Reading fluency is commonly identified through three properties of oral reading called..." 
q06_d1 <- "Accuracy, automaticity and semantics." 
q06_d2 <- "Accuracy, automaticity and flow."
q06_d3 <- "Accuracy, automaticity and syntax."
q06_c  <- "Accuracy, automaticity and prosody."

# q7
q07_q  <- "Which of these capabilities is *least* likely to contribute to a person's reading fluency?" 
q07_c  <- "Knowledge of comprehension strategies." 
q07_d1 <- "Word recognition."
q07_d2 <- "Syntactic knowledge."
q07_d3 <- "Vocabulary knowledge."

# q8
q08_q <- "Vocabulary tiers form a construct that is commonly used to prioritise particular words to teach. Tier 3 vocabulary can be thought of as…"
q08_d1 <- "Basic words that pupils are likely to learn through their everyday interactions."
q08_d2 <- "Words that are high frequency for mature language users and are found across a variety of domains."
q08_d3 <- "Words that are found primarily in scientific subjects."
q08_c  <- "Words that are quite low frequency and restricted to specific domains."

# q9
q09_q  <- "Which of these comprehension strategies has the *least* sound basis in evidence?"
q09_d1 <- "Self-questioning."
q09_c  <- "Extrapolating."
q09_d2 <- "Summarising."
q09_d3 <- "Clarifying."

# q10
q10_q  <- "In terms of reading, inferences are conclusions that..." 
q10_c  <- "Go beyond the literal meaning of the text."
q10_d1 <- "Don't require background knowledge." 
q10_d2 <- "Necessarily require conscious thinking." 
q10_d3 <- "Are not necessary when reading non-fiction texts."

# ---------- combine all questions and answers into data frame ----------

# grab matching elements and combine into lists
question <- lapply(ls(pattern="_q"), get)
correct <- lapply(ls(pattern="_c"), get)
distractor1 <- lapply(ls(pattern="_d1"), get)
distractor2 <- lapply(ls(pattern="_d2"), get)
distractor3 <- lapply(ls(pattern="_d3"), get)

# convert lists into dataframes
pretest <- do.call(rbind.data.frame, question)
names(pretest) <- "question"

tmp <- do.call(rbind.data.frame, correct)
names(tmp) <- "correct"
pretest$correct <- tmp$correct

tmp <- do.call(rbind.data.frame, distractor1)
names(tmp) <- "distractor1"
pretest$distractor1 <- tmp$distractor1

tmp <- do.call(rbind.data.frame, distractor2)
names(tmp) <- "distractor2"
pretest$distractor2 <- tmp$distractor2

tmp <- do.call(rbind.data.frame, distractor3)
names(tmp) <- "distractor3"
pretest$distractor3 <- tmp$distractor3

rm(tmp)
```


```{r, results='asis', echo=F}
knitr::kable(pretest, caption = "Pretest question with Correct Answers and Distractors")
```

```{r, include=FALSE, echo=F}
# ---------- randomise question and answers ----------

set.seed(1212) # make reproducible
q_rand <- pretest[sample(1:nrow(pretest)), ] # shuffle order of question
names(q_rand) <- c("question", "option1", "option2", "option3", "option4") # rename cols

# loop through all rows and shuffle answer options i.e., columns 2:ncol()
for (r in 1:nrow(q_rand)) {
  q_rand[r, 2:ncol(q_rand)] <-  q_rand[r,sample( 2:ncol(q_rand))]
  
}
```


```{r, results='asis', echo=F}
knitr::kable(q_rand, caption = "Pretest question Randomised")
```

```{r, include=F}
q_rand$order <- paste0("Q",1:nrow(q_rand))

order <- q_rand[, c("question", "order")]

pretest <- merge(pretest, order, by = "question")

for (c in 1:ncol(pretest)) {
  pretest[, c] <- gsub("[*]","", pretest[,c])
}

for (c in 1:ncol(q_rand)) {
  q_rand[, c] <- gsub("[*]","", q_rand[,c])
}

write.csv(pretest, "pretest_fin.csv", row.names = F)
write.csv(q_rand, "pretest_fin_rand.csv", row.names = F)
```

##### Question texts and answers

```{r, echo=FALSE, results='asis'}
for (q in 1:nrow(q_rand)) {
  
  # print to console/markdown
  cat("  \n\n")
  cat("  \n **", q, ". ", q_rand$question[q],"**", sep = "")
  cat("  \n\n (a) ", q_rand$option1[q])
  cat("  \n (b) ", q_rand$option2[q])
  cat("  \n (c) ", q_rand$option3[q])
  cat("  \n (d) ", q_rand$option4[q])
  

}
```

